Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-10-27 short.tex
additions to supplement

@@

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing
  primitives that explain an image. This set of drawing primitives is
  like a trace of the primitive commands issued by a graphics
  program. From this trace we use program synthesis techniques to
  recover a graphics program with constructs such as variable
  bindings, iterative loops, or simple kinds of conditionals. With a
  graphics program in hand, we can correct errors made by the deep
  network, cluster drawings by use of similar high-level geometric
  structures, and extrapolate drawings.  Taken together these results
  are a step towards agents that induce useful, human-readable
  programs from perceptual input.
\end{abstract}

@@

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{A token is an atomic symbol  -- so the model is \emph{not} predicting lines of code character-by-character, but instead knows about the syntactic structure of primitive commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
conditioned on the image features, then it emitted the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicted the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

Our network can ``derender'' random synthetic images
%those shown in Fig.~\ref{trainingData}.
by doing a beam search decoding to
recover  trace sets traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
drawing command), it has no way of recovering from the error.  
To derender an image with $n$ objects, it must correctly predict $n$
commands -- so its probability of success will decrease
exponentially in $n$, assuming it has a nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\citep{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate
\clearpage\begin{wrapfigure}{R}{8cm}
  \centering
      \includegraphics[width = 7cm]{figures/editDistance36.png}
%    \includegraphics[width = 4cm]{figures/accuracy.png}
      \caption{Parsing~\LaTeX~output after training on diagrams with $\leq 12$ objects. It generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
\end{wrapfigure}
\noindent
for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct.

@@

\subsection{Generalizing to hand drawings}\label{generalizingTheHandDrawings}
\begin{wrapfigure}{R}{6cm}
\centering  \begin{minipage}[t]{2.5cm}\centering\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}\\
   \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): hand drawing. (b): Rendering of the trace our model infers for (a). (c): noisy rendering of (b).}\label{handDrawingExamples}
\end{wrapfigure}

@@

We evaluate our system on 100 hand-drawn figures; see Fig.~\ref{drawingSuccesses}--\ref{drawingFailures}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.
For each drawing we annotated a ground truth trace and have the neurally guided SMC sampler
produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
We will show that the program synthesizer
corrects some of these small errors (Sec.~\ref{synthesizerHelpsParsing}).

@@

\begin{figure}[h]\centering
  \begin{minipage}[c]{0.57\textwidth} 
    \centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png}
  \end{minipage}\hfill%
      \begin{minipage}[c]{0.4\textwidth} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1,5,100)?
    Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 5 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth trace.}\label{drawingIntersectionOverUnion}
      \end{minipage}
            \vspace{-1cm}
\end{figure}

@@

\begin{table}[H]
  \caption{Grammar over graphics programs. We allow loops (\texttt{for}) with conditionals (\texttt{if}), vertical/horizontal reflections (\texttt{reflect}), variables (Var) and affine transformations ($\mathbb{Z}$\texttt{ * }Var\texttt{ + }$\mathbb{Z}$).}\label{DSL}
  \begin{tabular}{rl}\toprule
  Program$\to$&Statement; $\cdots$; Statement\\
  Statement$\to$&\texttt{circle}(Expression,Expression)\\
  Statement$\to$&\texttt{rectangle}(Expression,Expression,Expression,Expression)\\
  Statement$\to$&\texttt{line}(Expression,Expression,Expression,Expression,Boolean,Boolean)\\
  Statement$\to$&\texttt{for}$(0\leq \text{Var}  < \text{Expression})$\texttt{ \{ if }$(\text{Var} > 0)$\texttt{ \{ }Program\texttt{ \}; }Program\texttt{ \}}\\
  Statement$\to$&\texttt{reflect}$(\text{Axis})$\texttt{ \{ }Program\texttt{ \}}\\
  Expression$\to$&$\mathbb{Z}$\texttt{ * }Var\texttt{ + }$\mathbb{Z}$\\
  Var$\to$&A free (unused) variable\\
  $\mathbb{Z}\to$&an integer\\
  Axis$\to$&\texttt{X = }$\mathbb{Z}$ | \texttt{Y = }$\mathbb{Z}$\\\bottomrule
  \end{tabular}
\end{table}

@@

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%Describing Sketch's program synthesis
%algorithm is beyond the scope of this paper; see~\cite{solar2008program}.
At a high level, Sketch takes as input a space of programs, along with
a specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a minimum-cost
program satisfying the specification.  Sketch requires that the
program space be \emph{finite}, which here means that the depth of the
program syntax tree is bounded (we set the bound to 3),
but comes with the guarantee that it will
always eventually find the globally optimal (depth-bounded) solution.
In exchange for this optimality guarantee
it comes with no guarantees
on runtime.
For our domain synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
See Fig.~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving 58\% of problems within a minute.

@@

\begin{figure}[t]
%  \vspace{-0.6cm}
\centering  \begin{tabular}{m{1.5cm}llc}
    \toprule
    \textbf{Drawing}&\textbf{Trace}&\textbf{Program}&%\begin{tabular}{c}
      \textbf{Compression factor}%\\\textbf{factor}
      %      \end{tabular}\\
      \\
    \midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-29-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(2,15, 4,15)
Line(4,9, 4,13)
Line(3,11, 3,14)
Line(2,13, 2,15)
Line(3,14, 6,14)
Line(4,13, 8,13)
\end{lstlisting}
\end{minipage}&     \begin{minipage}{\exampleProgramSize} \begin{lstlisting}
for(i<3)
 line(i,-1*i+6,
      2*i+2,-1*i+6)
 line(i,-2*i+4,i,-1*i+6)
       \end{lstlisting}
     \end{minipage}&$\frac{6}{3} = 2\text{x}$\\\midrule
     \includegraphics[width = \exampleDrawingSize]{figures/expert-52-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(5,13, 2,10),
  arrow = True)
Circle(5,9)
Circle(8,5)
Line(2,8, 2,6), 
  arrow = True)
Circle(2,5)
\end{lstlisting}
\small\emph{... etc. ...; 13 lines}
\end{minipage}&
             \begin{minipage}{\exampleProgramSize}\begin{lstlisting}
circle(4,10)
for(i<3)
 circle(-3*i+7,5)
 circle(-3*i+7,1)
 line(-3*i+7,4,-3*i+7,2,
      arrow=True)
 line(4,9,-3*i+7,6,
      arrow=True)
\end{lstlisting}
\end{minipage}&$\frac{13}{6} = 2.2\text{x}$\\\midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-38-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(5,8)
Circle(2,8)
Circle(8,11)
Line(2,9, 2,10)
Circle(8,8)
Line(3,8, 4,8)
Line(3,11, 4,11)
\end{lstlisting}
  \small\emph{... etc. ...; 21 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<3)
 for(j<3)
  if(j>0)
   line(-3*j+8,-3*i+7,
        -3*j+9,-3*i+7)
  line(-3*i+7,-3*j+8,
       -3*i+7,-3*j+9)
  circle(-3*j+7,-3*i+7)
\end{lstlisting}
\end{minipage}&$\frac{21}{6} = 3.5\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-101-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Rectangle((1,10), (3,11))
Rectangle((1,12), (3,13))
Rectangle((4,8), (6,9))
Rectangle((4,10), (6,11))
\end{lstlisting}
  \small\emph{... etc. ...; 16 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<4)
 for(j<4)
  rectangle(-3*i+9,-2*j+6,
            -3*i+11,-2*j+7)
\end{lstlisting}
\end{minipage}&$\frac{16}{3} = 5.3\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-75-extra-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(11,14, 13,14), 
  arrow = True)
Circle(10,10)
Line(10,13, 10,11), 
  arrow = True)
Circle(6,10)
\end{lstlisting}
    \small\emph{... etc. ...; 15 lines}
\end{minipage}&
\begin{minipage}{\exampleProgramSize}
    \begin{lstlisting}
for(i<4)
 line(-4*i+13,4,-4*i+13,2,
      arrow=True)
 for(j<3)
  if(j>0)
   circle(-4*i+13,4*j+-3)
  line(-4*j+10,5,-4*j+12,5,
       arrow=True)
\end{lstlisting}
  \end{minipage}&$\frac{15}{6} = 2.5\text{x}$\\\midrule

@@

A search policy $\pi_\theta(\sigma  | T )$ takes as input a trace $T$ and predicts a distribution over synthesis problems, each of which is written $\sigma $ and corresponds to a set of possible programs to search over (so $\sigma \subset \text{DSL}$).
Good policies will prefer tractable program spaces,
so that the search procedure will terminate early, 
but should also prefer program spaces likely to contain
programs that concisely explain the data.
In general these two desiderata are in tension:
tractable synthesis problems involve searching over smaller spaces,
but smaller spaces are less likely to contain a minimum cost program
for explaining the trace.
Our goal now is to find the parameters of the policy, written $\theta$, which best navigate this trade-off.

@@

\textbf{Definition: Bias optimality.} A search algorithm is $n$-\emph{bias optimal}
with respect to a distribution $\probability_{\text{bias}}[\cdot ]$ if it is
guaranteed to find a solution in $\sigma $ after searching for at least time
$n\times\frac{t(\sigma )}{\probability_{\text{bias}}[\sigma ]}$, where $t(\sigma )$ is the time it
takes to verify that $\sigma $ contains a solution to the
search problem.

@@

We contrast our learned search policy
with two alternatives (see Fig.~\ref{policyEvaluation}):
(1) \emph{Sketch}, which poses the
entire problem wholesale to the Sketch program synthesizer;
and (2) an \emph{Oracle},
a policy which always picks the quickest to search $\sigma $
 also containing a minimum cost program.
Our approach improves upon Sketch by itself,
and comes close to the Oracle's performance.
One could never construct this Oracle,
because the agent does not know ahead of time which
$\sigma $'s contain minimum cost programs nor does it know how long each
$\sigma $ will take to search.
The Oracle is an upper bound on the performance of any search policy.
With this learned policy in hand we can synthesize 58\% of programs within a minute.
\begin{figure}[h]
  \centering  \includegraphics[width = 14cm]{figures/policyComparison.png}
  \caption{How long does it typically take to synthesize a minimum cost program? Sketch: out-of-the-box performance of the Sketch~\citep{solar2008program} program synthesizer. Oracle: baseline that picks the easiest synthesis problem containing a minimum cost program. Learned policy: a bias-optimal learned search policy running on an ideal timesharing machine. Oracle  upper bounds  the performance of any search policy. $\infty = $ timeout. Red dashed line is median time. Performance measured using cross validation.}\label{policyEvaluation}
\end{figure}

@@

We measure the similarity between two drawings by extracting features
of the lowest-cost programs that describe them. Our features are counts of the number of times that different components in the
DSL were used (Tbl.~\ref{DSL}).
We  then find drawings which are either close together or far apart in program feature space.
One could use many
alternative similarity metrics between drawings which would capture pixel-level similarities while missing high-level geometric similarities.
We used our learned distance metric between trace sets, $L_{\text{learned}}(\cdot|\cdot)$,
to find drawings that are either close together or far apart according to the learned
distance metric over images.
Tbl.~\ref{similarityMatrix} illustrates the kinds of drawings that these different metrics put closely together.

@@

\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we could change all of the circles to squares or make all of the lines be dashed.
We \textbf{extrapolate} figures
by increasing the number of times that loops are executed.
Extrapolating repetitive visuals patterns comes naturally to humans,
and building this ability into an application is practical:
imagine hand drawing a repetitive graphical model structure
and having our system automatically induce and extend the pattern.
Fig.~\ref{extrapolationFigure} shows extrapolations produced by our system;
see supplement for our full set of extrapolations.
\begin{figure}[H]
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix1.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix2.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix3.png}  
  \caption{Top, white: hand drawings. Bottom, black: extrapolations produced by
running loops for extra iterations.}\label{extrapolationFigure}
  \end{figure}
%\input{extrapolations.tex}

>>>>

\usepackage{amsthm}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

@@

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing
  primitives that explain an image. These drawing primitives are
  like a trace of the set of primitive commands issued by a graphics
  program. We learn a model that uses program synthesis techniques to
  recover a graphics program from that trace. These programs have constructs like variable
  bindings, iterative loops, or simple kinds of conditionals. With a
  graphics program in hand, we can correct errors made by the deep
  network, measure similarity between drawings by use of similar high-level geometric
  structures, and extrapolate drawings.  Taken together these results
  are a step towards agents that induce useful, human-readable
  programs from perceptual input.
\end{abstract}

@@

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{Tokens are atomic symbols in the programming language  -- so the model is \emph{not} predicting lines of code character-by-character, but instead the model knows about the syntactic structure of primitive drawing commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
conditioned on the image features, then it emitted the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicted the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

Our network can ``derender'' random synthetic images
%those shown in Fig.~\ref{trainingData}.
by doing a beam search decoding to
recover  trace sets maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
drawing command), it has no way of recovering from the error.  
To derender an image with $n$ objects, it must correctly predict $n$
commands -- so its probability of success will decrease
exponentially in $n$, assuming it has a nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\citep{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate
\clearpage\begin{wrapfigure}{R}{8cm}
  \centering
      \includegraphics[width = 7cm]{figures/editDistance36.png}
%    \includegraphics[width = 4cm]{figures/accuracy.png}
      \caption{Parsing~\LaTeX~output after training on diagrams with $\leq 12$ objects. It generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
\end{wrapfigure}
\noindent
for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct.

@@

\subsection{Generalizing to hand drawings}\label{generalizingTheHandDrawings}
\begin{wrapfigure}{R}{6cm}
\centering  \begin{minipage}[t]{2.5cm}\centering\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}\\
   \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): hand drawing. (b): Rendering of the trace set our model infers for (a). (c): noisy rendering of (b).}\label{handDrawingExamples}
\end{wrapfigure}

@@

We evaluate our system on 100 hand-drawn figures; see Fig.~\ref{drawingSuccesses}--\ref{drawingFailures}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.
For each drawing we annotated a ground truth trace set and have the neurally guided SMC sampler
produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
We will show that the program synthesizer
corrects some of these small errors (Sec.~\ref{synthesizerHelpsParsing}).

@@

\begin{figure}[h]\centering
  \begin{minipage}[c]{0.57\textwidth} 
    \centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png}
  \end{minipage}\hfill%
      \begin{minipage}[c]{0.4\textwidth} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1, 5, 100)?
    Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 5 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth trace. Red: the full model. Other colors: lesioned versions of our model.}\label{drawingIntersectionOverUnion}
      \end{minipage}
            \vspace{-1cm}
\end{figure}

@@

\begin{table}[H]
  \caption{Grammar over graphics programs. We allow loops (\texttt{for}) with conditionals (\texttt{if}), vertical/horizontal reflections (\texttt{reflect}), variables (Var) and affine transformations ($\mathbb{Z}\times$Var\texttt{+}$\mathbb{Z}$).}\label{DSL}
  \begin{tabular}{rl}\toprule
  Program$\to$&Statement; $\cdots$; Statement\\
  Statement$\to$&\texttt{circle}(Expression,Expression)\\
  Statement$\to$&\texttt{rectangle}(Expression,Expression,Expression,Expression)\\
  Statement$\to$&\texttt{line}(Expression,Expression,Expression,Expression,Boolean,Boolean)\\
  Statement$\to$&\texttt{for}$(0\leq \text{Var}  < \text{Expression})$\texttt{ \{ if }$(\text{Var} > 0)$\texttt{ \{ }Program\texttt{ \}; }Program\texttt{ \}}\\
  Statement$\to$&\texttt{reflect}$(\text{Axis})$\texttt{ \{ }Program\texttt{ \}}\\
  Expression$\to$&$\mathbb{Z}\times$Var\texttt{+}$\mathbb{Z}$\\
  Var$\to$&A free (unused) variable\\
  $\mathbb{Z}\to$&an integer\\
  Axis$\to$&\texttt{X = }$\mathbb{Z}$ | \texttt{Y = }$\mathbb{Z}$\\\bottomrule
  \end{tabular}
\end{table}

@@

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%Describing Sketch's program synthesis
%algorithm is beyond the scope of this paper; see~\cite{solar2008program}.
At a high level, Sketch takes as input a space of programs, along with
a specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a minimum-cost
program satisfying the specification.  Sketch requires a
 \emph{finite program space}, which here means that the depth of the
program syntax tree is bounded (we set the bound to 3),
but comes with the guarantee that it will
always eventually find the globally optimal (depth-bounded) solution.
In exchange for this optimality guarantee
it comes with no guarantees
on runtime.
For our domain synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
See Fig.~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving 58\% of problems within a minute.

@@

\begin{figure}[t]
%  \vspace{-0.6cm}
\centering  \begin{tabular}{m{1.5cm}llc}
    \toprule
    \textbf{Drawing}&\textbf{Trace Set}&\textbf{Program}&%\begin{tabular}{c}
      \textbf{Compression factor}%\\\textbf{factor}
      %      \end{tabular}\\
      \\
    \midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-29-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(2,15, 4,15)
Line(4,9, 4,13)
Line(3,11, 3,14)
Line(2,13, 2,15)
Line(3,14, 6,14)
Line(4,13, 8,13)
\end{lstlisting}
\end{minipage}&     \begin{minipage}{\exampleProgramSize} \begin{lstlisting}
for(i<3)
 line(i,-1*i+6,
      2*i+2,-1*i+6)
 line(i,-2*i+4,i,-1*i+6)
       \end{lstlisting}
     \end{minipage}&$\frac{6}{3} = 2\text{x}$\\\midrule
     \includegraphics[width = \exampleDrawingSize]{figures/expert-52-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(5,13, 2,10),
  arrow = True)
Circle(5,9)
Circle(8,5)
Line(2,8, 2,6), 
  arrow = True)
Circle(2,5)
\end{lstlisting}
\small\emph{... etc. ...; 13 lines}
\end{minipage}&
             \begin{minipage}{\exampleProgramSize}\begin{lstlisting}
circle(4,10)
for(i<3)
 circle(-3*i+7,5)
 circle(-3*i+7,1)
 line(-3*i+7,4,-3*i+7,2,
      arrow=True)
 line(4,9,-3*i+7,6,
      arrow=True)
\end{lstlisting}
\end{minipage}&$\frac{13}{6} = 2.2\text{x}$\\\midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-38-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(5,8)
Circle(2,8)
Circle(8,11)
Line(2,9, 2,10)
Circle(8,8)
Line(3,8, 4,8)
Line(3,11, 4,11)
\end{lstlisting}
  \small\emph{... etc. ...; 21 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<3)
 for(j<3)
  if(j>0)
   line(-3*j+8,-3*i+7,
        -3*j+9,-3*i+7)
  line(-3*i+7,-3*j+8,
       -3*i+7,-3*j+9)
  circle(-3*j+7,-3*i+7)
\end{lstlisting}
\end{minipage}&$\frac{21}{6} = 3.5\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-101-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Rectangle((1,10), (3,11))
Rectangle((1,12), (3,13))
Rectangle((4,8), (6,9))
Rectangle((4,10), (6,11))
\end{lstlisting}
  \small\emph{... etc. ...; 16 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<4)
 for(j<4)
  rectangle(-3*i+9,-2*j+6,
            -3*i+11,-2*j+7)
\end{lstlisting}
\end{minipage}&$\frac{16}{3} = 5.3\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-75-extra-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(11,14, 13,14), 
  arrow = True)
Circle(10,10)
Line(10,13, 10,11), 
  arrow = True)
Circle(6,10)
\end{lstlisting}
    \small\emph{... etc. ...; 15 lines}
\end{minipage}&
\begin{minipage}{\exampleProgramSize}
    \begin{lstlisting}
for(i<4)
 line(-4*i+13,4,-4*i+13,2,
      arrow=True)
 for(j<3)
  if(j>0)
   circle(-4*i+13,4*j+-3)
  line(-4*j+10,5,-4*j+12,5,
       arrow=True)
\end{lstlisting}
  \end{minipage}&$\frac{15}{6} = 2.5\text{x}$\\\midrule

@@

A search policy $\pi_\theta(\sigma  | T )$ takes as input a trace $T$ and predicts a distribution over synthesis problems, each of which is written $\sigma $ and corresponds to a set of possible programs to search over (so $\sigma \subset \text{DSL}$).
Good policies will prefer tractable program spaces,
so that the search procedure will terminate early, 
but should also prefer program spaces likely to contain
programs that concisely explain the data.
These two desiderata are in tension:
tractable synthesis problems involve searching over smaller spaces,
but smaller spaces are less likely to contain a minimum cost program
for explaining the trace.
Our goal now is to find the parameters of the policy, written $\theta$, which best navigate this trade-off.

@@

\noindent\textbf{Definition: Bias-optimality.}
   A search algorithm is $n$-\emph{bias optimal}
with respect to a distribution $\probability_{\text{bias}}[\cdot ]$ if it is
guaranteed to find a solution in $\sigma $ after searching for at least time
$n\times\frac{t(\sigma )}{\probability_{\text{bias}}[\sigma ]}$, where $t(\sigma )$ is the time it
takes to verify that $\sigma $ contains a solution to the
search problem.

@@

We contrast our learned search policy
with two alternatives (see Fig.~\ref{policyEvaluation}):
(1) \emph{Sketch}, which poses the
entire problem wholesale to the Sketch program synthesizer;
and (2) an \emph{Oracle},
a policy which always picks the quickest to search $\sigma $
 also containing a minimum cost program.
Our approach improves upon Sketch by itself,
and comes close to the Oracle's performance.
One could never construct this Oracle,
because the agent does not know ahead of time which
$\sigma $'s contain minimum cost programs nor does it know how long each
$\sigma $ will take to search.
The Oracle is an upper bound on the performance of any search policy.
With this learned policy in hand we can synthesize 58\% of programs within a minute.
\begin{figure}[h]
  \centering  \includegraphics[width = 14cm]{figures/policyComparison.png}
  \caption{How long does it typically take to synthesize a minimum cost program? Sketch: out-of-the-box performance of the Sketch~\citep{solar2008program} program synthesizer. Oracle: baseline that picks the easiest synthesis problem containing a minimum cost program. Learned policy: a bias-optimal learned search policy running on an ideal timesharing machine. Oracle  upper bounds  the performance of any search policy. $\infty = $ timeout. Red dashed line is median time. Learned policy evaluated using 20-fold cross validation.}\label{policyEvaluation}
\end{figure}

@@

We measure the similarity between two drawings by extracting features
of the lowest-cost programs that describe them. Our features are counts of the number of times that different components in the
DSL were used (Tbl.~\ref{DSL}).
We  then find drawings which are either close together or far apart in program feature space.
One could use many
alternative similarity metrics between drawings which would capture pixel-level similarities while missing high-level geometric similarities.
We used our learned distance metric between trace sets, $L_{\text{learned}}(\cdot|\cdot)$,
to find drawings that are either close together or far apart according to the learned
distance metric over images.
Tbl.~\ref{similarityMatrix} illustrates the kinds of drawings that these different metrics put closely together,
while the supplement shows low dimensional projections of
program and image features.

@@

\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we could change all of the circles to squares or make all of the lines be dashed.
We \textbf{extrapolate} figures
by increasing the number of times that loops are executed.
Extrapolating repetitive visuals patterns comes naturally to humans,
and building this ability into an application is practical:
imagine hand drawing a repetitive graphical model structure
and having our system automatically induce and extend the pattern.
Fig.~\ref{extrapolationFigure} shows extrapolations produced by our system.
\begin{figure}[H]
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix1.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix2.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix3.png}  
  \caption{Top, white: hand drawings. Bottom, black: extrapolations produced by
running loops for extra iterations.}\label{extrapolationFigure}
  \end{figure}
%\input{extrapolations.tex}
